---
title: "Cycle Stats EDA"
author: "Will Hore-Lacy"
date: "04/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
library(leaflet)
library(sf)
library(knitr)
library(readxl)
```

## VicRoads Bicycle Counters

```{r, fig.width=10, fig.height=7}

bikeSites <- as.data.table(read_xlsx("./data/VicRoads_Bike_Site_Number_Listing.xlsx"))

bikeSites <- bikeSites[!is.na(LAT)]
bikeSites[,7:13:=NULL]
setnames(bikeSites,"LOC DESC","LOC_DESC")

leaflet(bikeSites) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addCircleMarkers(lng=~LONG, lat=~LAT, radius=4, label=~paste0(SITE_XN_ROUTE,": ",LOC_DESC)) %>%
    addCircleMarkers(data=bikeSites[SITE_XN_ROUTE==6415], lng=~LONG, lat=~LAT,
                     label=~paste0(SITE_XN_ROUTE,": ",LOC_DESC), 
                     radius=5, color="red", opacity=.9)



```

Try examining 6415 site on the Main Yarra Trail: `r bikeSites[SITE_XN_ROUTE==6415, LOC_DESC]`

```{r site59485}
MYTid <- 6415
# MYTsiteId <- bikeSites[LOC_LEG==59485, SITE_XN_ROUTE] # This is the id used in the file names

```

```{r unzip, eval=TRUE}

MYT <- data.table()

for (zipFile in list.files(file.path("./data","Bicycle_Volume_Speed_2018"), pattern=".zip$", full.names = T)) {
    
    # get list of csv files in the zip file
    csvFiles <- as.data.table(unzip(zipFile,list=T))
    
    # if id is found in file names, extract just that file to working dir
    if(sum(grepl(MYTid,csvFiles$Name))>0) {
        
        csvFiles[grepl(MYTid,Name),Name]
        MYT <- rbind(MYT,
                     as.data.table(
                         read.table(
                             unzip(zipFile, files=csvFiles[grepl(MYTid,Name),Name]), 
                             header=T, sep=",")))
        
    }
    
}

# MYT[,AMPM:= ]

ggplot(MYT, aes(x=SPEED)) +
    stat_density(geom="line") +
    facet_grid(DIRECTION ~.) +
    labs(x="Speed (km/h)")

```

```{r}
Sys.setenv(TZ='GMT')

convTime <- function(x) {

    require(chron)
    x1 <- as.character(x)
    return(chron(times=x1))
    
}

MYT[,TIME2:=convTime(TIME)]

ggplot(MYT, aes(x=TIME2)) +
    stat_density(geom="line") +
    facet_grid(DIRECTION ~.) +
    scale_x_chron(format="%H:%M") +
    labs(x="Time of day")



```

First graph data created with a DT group by, may not be correct.

```{r}
# create one hour time bins and find max density

# Unsure if this is the correct aggregation
MYThour <- MYT[,.(count=length(DATA_TYPE)/length(unique(MYT$DATE))), by=.(hours(TIME2),DIRECTION)]
# length(unique(MYT$DATE))

MYThour[,maximum:=0]
MYThour[count==max(MYThour[DIRECTION=="E",count]),maximum:=1]
MYThour[count==max(MYThour[DIRECTION=="W",count]),maximum:=1]

ggplot(MYThour, aes(x=hours, y=count, fill=as.factor(maximum))) +
    geom_bar(stat="identity") +
    facet_grid(DIRECTION ~.) +
    labs(x="Time of day")


# Try another
MYT[,time_hr:=hours(TIME2)]
MYT[,DATE:=as.Date(as.character(DATE), format="%d/%m/%Y")]

MYTcast1 <- dcast(MYT, time_hr + DATE + DIRECTION ~., fun.aggregate = length)
setnames(MYTcast1,".","count")

MYTcast1 <- dcast(MYTcast1, time_hr + DIRECTION ~., fun.aggregate = mean, value.var = "count")
setnames(MYTcast1,".","count")

MYTcast1[,maximum:=0]
MYTcast1[count==max(MYTcast1[DIRECTION=="E",count]),maximum:=1]
MYTcast1[count==max(MYTcast1[DIRECTION=="W",count]),maximum:=1]

ggplot(MYTcast1, aes(x=time_hr, y=count, fill=as.factor(maximum))) +
    geom_bar(stat="identity") +
    facet_grid(DIRECTION ~.) +
    labs(x="Time of day")

```

Inerestingly data is only for weekdays.

```{r}

MYTcast2 <- dcast(MYT, time_hr + DATE + DIRECTION ~., fun.aggregate = length)
setnames(MYTcast2,".","count")

# allocate to week or weekend 
MYTcast2[, wkDay := weekdays(MYTcast2$DATE[1:5])]
MYTcast2[, weekend := "Weekday"]
MYTcast2[wkDay == "Saturday" | wkDay == "Sunday", weekend := "Weekend"]

MYTcast2 <- dcast(MYTcast2, time_hr + DIRECTION + weekend ~., fun.aggregate = mean, value.var = "count")
setnames(MYTcast2,".","count")

# MYTcast1[,maximum:=0]
# MYTcast1[count==max(MYTcast1[DIRECTION=="E",count]),maximum:=1]
# MYTcast1[count==max(MYTcast1[DIRECTION=="W",count]),maximum:=1]

ggplot(MYTcast2, aes(x=time_hr, y=count)) +
    geom_bar(stat="identity") +
    facet_grid(DIRECTION ~ weekend) +
    labs(x="Time of day")


```

Looks like I'm missing data from a few months.

```{r fig.width=10, fig.height=10}

MYT[, mnth := months(DATE) ]

MYTcast3 <- dcast(MYT, time_hr + DATE + DIRECTION + mnth ~., fun.aggregate = length)
setnames(MYTcast3,".","count")

MYTcast3 <- dcast(MYTcast3, time_hr + DIRECTION + mnth ~., fun.aggregate = mean, value.var = "count")
setnames(MYTcast3,".","count")
MYTcast3[,mnth := factor(mnth, 
                        levels=c("January", "February", "March", "April", "May", "June", 
                                 "July", "August", "September", "October", "November", "December"),
                        ordered = T)]


ggplot(MYTcast3, aes(x=time_hr, y=count)) +
    geom_bar(stat="identity") +
    facet_grid( mnth ~ DIRECTION ) +
    labs(x="Time of day")
```

```{r}

ggplot(MYT, aes(x=DATE)) +
    stat_density(geom="line", adjust = .4) +
    facet_grid(DIRECTION ~.) +
    labs(x="Speed (km/h)")


```



## MCC Pedestrian Counters

```{r}

pedSites <- fread("./data/Pedestrian_sensor_locations.csv")

# pedSites <- st_as_sf(pedSites,coords = c("Longitude", "Latitude"))


leaflet(pedSites) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addCircleMarkers(lng=~Longitude, lat=~Latitude, radius=4)

```

## Pedestrian Volume Data

```{r}

pedDT <- fread("./data/Pedestrian_volume__updated_monthly_.csv")

head(pedDT)

summary(pedDT)

```

Look at the "Southbank" sensor. Aggregate to days of the week and hours.

```{r}

southbank <- pedDT[Sensor_Name=="Southbank"]

southbank2 <- dcast(southbank, Day + Time ~ "Hourly_Counts", value.var = "Hourly_Counts", fun.aggregate = mean)
names(southbank2)[3] <- "Hourly_Counts"
southbank2[,Day:=factor(Day,levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))]

ggplot(southbank2, aes(x=Time,y=Hourly_Counts,colour=Day)) +
    geom_line()

```
